---
title: "TMEM and BrCA Metastasis: Reproducible Analysis Code"
author: "Chenxin Zhang"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, warning=FALSE}
rm(list = ls())

knitr::opts_chunk$set(
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  include = FALSE,
  dpi = 300,
  comment = NA
)

set.seed(820)

library(tidyverse)
library(readxl)
library(openxlsx)
library(survminer)
library(survival)
library(gtsummary)
library(flextable)
library(officer)
library(stringr)
library(purrr)
library(broom)
library(janitor)

select <- dplyr::select

code_dir <- if (!is.null(knitr::current_input())) {
  dirname(normalizePath(knitr::current_input(), mustWork = TRUE))
} else if (file.exists("TMEM_and_BrCA_Metastasis_analysis.Rmd")) {
  getwd()
} else if (dir.exists(file.path(getwd(), "TMEM CODE"))) {
  file.path(getwd(), "TMEM CODE")
} else {
  getwd()
}

project_dir <- normalizePath(file.path(code_dir, ".."), mustWork = TRUE)
data_dir <- if (dir.exists(file.path(code_dir, "data"))) {
  file.path(code_dir, "data")
} else {
  file.path(project_dir, "data")
}
results_dir <- file.path(code_dir, "results")
km_dir <- file.path(results_dir, "km_curves")
template_docx <- file.path(code_dir, "landscape_example.docx")

dir.create(results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(km_dir, showWarnings = FALSE, recursive = TRUE)

tertile_markers <- c("tmem_tertile", "mena_calc_tertile", "mena_inv_tertile")
main_markers <- c(
  "tmem_tertile", "tmem_score",
  "mena_calc_tertile", "mena_calc_score",
  "mena_inv_tertile", "mena_inv_score"
)
numeric_mutual_markers <- c(
  "tmem_score + mena_calc_score",
  "tmem_score + mena_inv_score",
  "tmem_score + mena_calc_score + mena_inv_score",
  "mena_calc_score + mena_inv_score"
)
tertile_mutual_markers <- c(
  "tmem_tertile + mena_calc_tertile",
  "tmem_tertile + mena_inv_tertile",
  "tmem_tertile + mena_calc_tertile + mena_inv_tertile",
  "mena_calc_tertile + mena_inv_tertile"
)

missing_values <- c(
  "999-Missing", "9-Missing", "99-Missing", "NULL",
  "Unable to be Scored", "5-Unknown/Unstaged", "3-Uknown/Indeterminate",
  "8-UNK", "8=Patient Unavailable", "7=Patient Declined", "3=Unknown",
  "6-Results not found", "5-Not documented", "4-Unk", "5-No info",
  "5-No Info", "Unknown", 9999, 999
)

format_p <- function(x) {
  ifelse(
    is.na(x), "",
    ifelse(
      x < 0.001, "<0.001",
      ifelse(x < 0.01, sprintf("%.3f", x), sprintf("%.2f", x))
    )
  )
}

format_p_table1 <- function(x) {
  ifelse(
    is.na(x), "",
    ifelse(x < 0.001, "<0.001", sprintf("%.4f", x))
  )
}

# Race uses Monte Carlo chi-square p-values in the baseline tables.
race_chisq_test_args <- list(
  race2 = list(simulate.p.value = TRUE)
)

new_results_doc <- function() {
  if (file.exists(template_docx)) {
    read_docx(template_docx)
  } else {
    read_docx()
  }
}

```


# Import and clean the raw data
```{r}
rawdf = read_excel(file.path(data_dir, "NSABP Master File.xlsx"), sheet = "Chart Review", na = missing_values) %>%
  clean_names() %>%
  rename(id = bank_id) 

master = read_excel(file.path(data_dir, "NSABP Master File.xlsx"), sheet = "Master with Scores", na = missing_values) %>%
  clean_names() %>%
  rename(id = bank_id) 

master <- master %>%
  mutate(across(
    where(is.character),
    ~ str_replace(as.character(.), "^\\d+[\\-=]", "")
  ))

# Import HER2 restain data.
her_df = read_excel(file.path(data_dir, "NSABP Master File.xlsx"), sheet = "97 HER2 Restains") %>%
  mutate(her2_tpc = ifelse(`HER2 Result: TPC LAB` %in% c("Tissue Fell Off", "Lymph Node"), NA, `HER2 Result: TPC LAB`)) %>%
  rename(id = `Sample Number`) %>%
  select(id, her2_tpc)

# Join chart review, score, and HER2 data.
df = rawdf %>%
  left_join(master %>% select(-usi), by = "id") %>%
  left_join(her_df, by = "id")


# Calculate MenaCalc score.
zscore <- function(x) {
   (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
 }
df = df %>% mutate(
  mena11a_z = zscore(cy5_mean_intensity_mena11a),
  pan_mena_z = zscore(tritc_mean_intensity_pan_mena),
  mena_calc_score = pan_mena_z - mena11a_z) %>% select(-mena11a_z,-pan_mena_z)

# Analytic cohort: 814 eligible patients with tumor block received and
# successful TMEM staining.
df = df %>%
  filter(
    a_eligible == "Eligible",
    b_block_received == "Block Received",
    c_tmem_stain_status %in% c(
      "TMEM Stained Successfully, Score 0",
      "TMEM Stained Successfully"
    )
  ) %>%
  select(
    -a_eligible, -b_block_received, -c_tmem_stain_status, -protocol, -seq_no,
    -mena_inv_exclusion_remark, -mena_calc_exlusion_remark,
    -cytokeratin_mask_area, -cy5_mean_intensity_mena11a,
    -tritc_mean_intensity_pan_mena
  )
```

## Clean data
```{r}
df = df  %>% mutate(
  
  mpsiz = ifelse(mpsiz ==99.9,NA,mpsiz),
  trt = factor(trt, levels = c(1,2), labels = c("AC", "AC + PTX")), 
  
  mense = ifelse(mense == 9, NA,mense),
  mense = factor(mense, levels = c(1:3), labels = c("Pre-menopausal", "Currently experiencing","Post-menopausal")),#no unknown value
  
  tmast = factor(tmast, levels = c(1,2), labels = c("Lumpectomy + Axillary Dissection", "Mastectomy")),
  
  hr = factor(ifelse(er==1 & pr==1, 1,2), levels = c(1,2), labels = c("Negative", "Positive")),
  
  er = factor(er, levels = c(1,2), labels = c("Negative", "Positive")),#no "Equivocal" or "Unknown" categories
  
  pr = factor(pr, levels = c(1,2), labels = c("Negative", "Positive")),#no "Equivocal", "Unknown"
  
  her2_gene_expression = ifelse(her2_3<11.5, "Negative",ifelse(her2_3>=11.5, "Positive",NA)),
  #combine her2_tpc and her2_gene_expression
  `her2_tpc_onco` = ifelse(is.na(her2_tpc),her2_gene_expression,her2_tpc),
  
  er_her2 = factor(case_when(
    `her2_tpc_onco` == "Negative" & er == "Negative" ~ "her2-er-",
    `her2_tpc_onco` == "Positive" & er == "Negative" ~ "her2+er-",
    `her2_tpc_onco` == "Negative" & er == "Positive" ~ "her2-er+",
    `her2_tpc_onco` == "Positive" & er == "Positive" ~ "her2+er+",
    TRUE ~ "Missing at least one"
  ), levels = c("her2-er-","her2-er+","her2+er-","her2+er+","Missing at least one")),
  
  tam = factor(tam, levels = c(0,1), labels = c("N", "Y")),
  
  grade = ifelse(grade == 9, NA, grade),
  grade = factor(
    grade,
    levels = c(1:3),
    labels = c(
      "Well differentiated",
      "Moderately differentiated",
      "Poorly differentiated"
    )
  ),

  race2 = case_when(
    race == 1 ~ 1,
    race == 2 ~ 2,
    race == 3 ~ 3,
    race == 4 ~ 4,
    race == 5 ~ 5,
    race == 6 ~ 5,
    race == 7 ~ 4,
    race == 8 ~ 4,
    race == 9 ~ 5,
    race == 10 ~ 5
  ),
  race2 = factor(race2, levels = c(1:5),labels = c("NH-WT", "Hispanic","Black", "Asian", "Other/NA")),
  
  mena_inv_score=as.numeric(mena_inv_score),
  mena_calc_score=as.numeric(mena_calc_score))


```

## Create survival time and event variables
```{r}

# OS: Overall survival is defined as the time from curative surgery to death from
# any cause. The OS variables are tos (time) and os (event).

# DFS: Disease-free survival events include local, regional, or distant recurrence;
# contralateral breast cancer; non-breast second primary cancers; and death before
# recurrence. Patients are censored at the end of follow-up
# or at loss to follow-up.


# DDFS: Distant disease-free survival focuses on distant metastasis. Events other
# than distant recurrence are treated as censoring events.

# Variables ending in "_5ys" indicate outcomes administratively truncated at 5 years.
# For these variables, follow-up time is capped at 60 months; events occurring after
# 60 months are treated as censored at 60 months.

df = df %>% mutate(
  # Outcome times are in months.
  locrec = as.numeric(locrec),
  tlocrec = as.numeric(tlocrec),
  regrec = as.numeric(regrec),
  treg_rec = as.numeric(treg_rec),
  disrec = as.numeric(disrec),
  tdis_rec = as.numeric(tdis_rec),
  secpri = as.numeric(secpri),
  tp2 = as.numeric(tp2),
  death = as.numeric(death),
  tlfup = as.numeric(tlfup)
)

# Derive full-follow-up and 5-year truncated endpoints.
df = df %>% rowwise() %>% 
   mutate(
  #OS
     os = death,
    tos = tlfup,
    
   os_5ys = ifelse(tos > 60 , 0, os),
   tos_5ys = ifelse(tos > 60 ,60 , tos), 
   
  #DDFS
  ddfs = disrec,
  tddfs = ifelse(ddfs==1, tdis_rec, tlfup),
  
  ddfs_5ys = ifelse(tddfs > 60, 0, ddfs),
  tddfs_5ys = ifelse(tddfs > 60, 60, tddfs),
  
  #DFS
    dfs = ifelse(locrec==0 & regrec==0 & disrec==0 & secpri==0 & death==0, 0,1),
    tdfs = min(tlocrec, treg_rec, tdis_rec, tp2, tlfup, na.rm = T),
    
    dfs_5ys = ifelse(tdfs > 60 , 0, dfs),
    tdfs_5ys = ifelse(tdfs > 60 ,60 , tdfs)
    
    ) %>% ungroup()
```

## Categorize biomarkers into tertiles
```{r}
# Cutoffs are derived from all eligible patients.
quantile(df$tmem_score, probs = c(0,1/3,2/3,1), na.rm = TRUE)
quantile(df$mena_inv_score, probs = c(0,1/3,2/3,1), na.rm = TRUE)
quantile(df$mena_calc_score, probs = c(0,1/3,2/3,1), na.rm = TRUE)

df = df %>% mutate(
  tmem_tertile = factor(case_when(
    tmem_score < 1.169908 ~ "Low",
    tmem_score >= 1.169908 & tmem_score < 10.408976 ~ "Mid",
    tmem_score >= 10.408976~ "High",
    TRUE ~ NA
  ), levels = c("Low", "Mid","High")),
  
  mena_inv_tertile = factor(case_when(
    mena_inv_score < 13.82770 ~ "Low",
    mena_inv_score >= 13.82770 & mena_inv_score < 18.16060 ~ "Mid",
    mena_inv_score >= 18.16060~ "High",
    TRUE ~ NA
  ), levels = c("Low", "Mid","High")),
  
  mena_calc_tertile = factor(case_when(
    mena_calc_score < -0.11642614 ~ "Low",
    mena_calc_score >= -0.11642614 & mena_calc_score < 0.07681679 ~ "Mid",
    mena_calc_score >= 0.07681679~ "High",
    TRUE ~ NA
  ), levels = c("Low", "Mid","High"))
)

df = df %>%
  mutate(
    tmem_num = as.numeric(tmem_tertile),
    tmem_ord = factor(tmem_tertile,
                              levels = c("Low", "Mid", "High"),
                              ordered = TRUE),
    
    mena_inv_num = as.numeric(mena_inv_tertile),
    mena_inv_ord = factor(mena_inv_tertile,
                                  levels = c("Low", "Mid", "High"),
                                  ordered = TRUE),
    mena_calc_num = as.numeric(mena_calc_tertile),
    mena_calc_ord = factor(mena_calc_tertile,
                                   levels = c("Low", "Mid", "High"),
                                   ordered = TRUE)
  )
```

## Final analysis datasets
```{r}
# Create df2 for Cox regression models.
# In df2, missing tumor size and recurrence score are imputed with their
# full-cohort medians, and missing tumor grade and HER2 status are retained as
# explicit "Unknown" categories so that patients with missing covariate values
# are not excluded from adjusted models.
# The original df is retained for descriptive tables and unmodified summaries.

df2 <- df %>%
  mutate(
    # replace missing tumor size with median
    mpsiz = ifelse(is.na(mpsiz), median(mpsiz, na.rm = TRUE), mpsiz),

    # replace missing recurrence score with the full-cohort median
    recurrence = ifelse(
      is.na(recurrence),
      median(recurrence, na.rm = TRUE),
      recurrence
    ),

    # Recode missing tumor grade as a separate group.
    grade = factor(
      ifelse(is.na(grade), "unknown", grade),
      labels = c(
        "Well differentiated",
        "Moderately differentiated",
        "Poorly differentiated",
        "Unknown"
      )
    ),
       
    `her2_tpc_onco` = factor(ifelse(is.na(`her2_tpc_onco`), "unknown",`her2_tpc_onco`),labels = c("Negative","Positive","Unknown"))
    
)


# Define the ER+/HER2- subgroup in both the original dataset and the modeling
# dataset. df_sub is used for descriptive summaries, and df2_sub is used for
# adjusted Cox models.

df_sub = df %>% filter(er_her2 == "her2-er+")
df2_sub = df2 %>% filter(er_her2 == "her2-er+")
```


## Median follow-up of overall survival
```{r}

# Reverse Kaplan-Meier for follow-up time.
fit_followup <- survfit(
  Surv(tos, 1 - os) ~ 1,
  data = df
)

q <- quantile(fit_followup, probs = c(0.25, 0.5, 0.75))$quantile
q_years <- q / 12

cat(
  sprintf(
    "Median [IQR] follow-up: %.2f [%.2f–%.2f] years",
    q_years[2],
    q_years[1],
    q_years[3]
  )
)#Median [IQR] follow-up: 11.60 [11.04–12.31] years
```

# Mean values of biomarkers of interest
```{r}
biomarker_summary <- tibble(
  marker = c("TMEM doorway score", "MenaINV score", "MenaCalc score"),
  variable = c("tmem_score", "mena_inv_score", "mena_calc_score")
) %>%
  mutate(
    mean = map_dbl(variable, ~ mean(df[[.x]], na.rm = TRUE)),
    sd = map_dbl(variable, ~ sd(df[[.x]], na.rm = TRUE)),
    summary = sprintf("%s: %.4f (%.4f)", marker, mean, sd)
  )

cat(paste(biomarker_summary$summary, collapse = "\n"))
```

# Cox Model
Clinical variables to consider are:
(a)  Age
(b)  Tumor grade
(c)  No. of lymph nodes
(d)  Estrogen receptor status
(e)  Tumor size
(f)  Type of surgery
(g)  HER2 gene expression
(h)  Randomization (treatment) arm. 
## Define all outcomes
```{r}
mycovars = "+ age + grade + pnod + er + mpsiz + tmast + her2_tpc_onco + trt"
subcovars = "+ age + grade + pnod + mpsiz + tmast + trt"

outcomes <- list(
  DDFS = list(
    time = "tddfs",
    event = "ddfs",
    outcome_label = "DDFS"
  ),
  DDFS_5YS = list(
    time = "tddfs_5ys",
    event = "ddfs_5ys",
    outcome_label = "DDFS (5-year)"
  ),
  DFS = list(
    time = "tdfs",
    event = "dfs",
    outcome_label = "DFS"
  ),
  DFS_5YS = list(
    time = "tdfs_5ys",
    event = "dfs_5ys",
    outcome_label = "DFS (5-year)"
  )
)

mylabel = list(
      age               ~ "Age",
      grade     ~ "Tumor grade",
      pnod              ~ "Positive lymph nodes",
      mpsiz             ~ "Tumor size (cm)",
      tmast             ~ "Type of surgery",
      trt               ~ "Treatment"
    )
```

## Build Cox model table functions
```{r}
# Centralized data switch for optional time-varying models.
get_cox_data <- function(outcome_name, interaction, data) {
  if (interaction != "time_vary") return(data)

  switch(
    outcome_name,
    OS       = cox_os,
    DDFS     = cox_ddfs,
    DFS      = cox_dfs,
    OS_5YS   = cox_os_5ys,
    DDFS_5YS = cox_ddfs_5ys,
    DFS_5YS  = cox_dfs_5ys,
    data
  )
}

build_cox_formula <- function(time, event, mymarker, interaction, covars) {
  as.formula(
    paste0(
      "Surv(", time, ", ", event, ") ~ ",
      mymarker,
      if (interaction == "time_vary") ":strata(tgroup)" else "",
      covars
    )
  )
}

make_cox_table <- function(data, mymarker, covars,interaction = "NO") {

  tbls <- purrr::imap(
    outcomes,
    function(outcome, outcome_name) {

      f <- build_cox_formula(
        time  = outcome$time,
        event = outcome$event,
        mymarker = mymarker,
        interaction = interaction,
        covars = covars
      )

      d <- get_cox_data(outcome_name, interaction, data)
      
      fit <- withCallingHandlers(
        coxph(f, data = d, ties = "breslow", model = TRUE, x = TRUE, y = TRUE),
        warning = function(w) {
          message("\nCox warning in outcome: ", outcome_name,
                  "  [", outcome$outcome_label, "]")
          message("Formula: ", deparse(f))
          message("Warning: ", conditionMessage(w), "\n")
          invokeRestart("muffleWarning")
        }
      )
 
      tbl_regression(
        fit,
        label = mylabel,
        exponentiate = TRUE,
        pvalue_fun = format_p
      ) %>%
        add_global_p(keep = TRUE) %>%
        bold_p(t = 0.05) %>%
        bold_labels() %>%
        italicize_levels()
    }
  )

  ft = tbl_merge(
    tbls = tbls,
    tab_spanner = paste0("**", purrr::map_chr(outcomes, "outcome_label"), "**")
  ) 
  ft <- ft %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(order_flag = if_else(grepl(":", variable), 0, 1)) %>%
      dplyr::arrange(order_flag) %>%
      dplyr::select(-order_flag)
  )
  
  ft = ft %>%
    as_flex_table() %>%
    set_table_properties(layout = "autofit") %>%
    padding(
      padding.top = 0, padding.bottom = 0,
      padding.left = 1, padding.right = 1
    ) %>% autofit() %>% vline(j = c(4,7,10), border = fp_border(color = "black", width = 1))
  # Highlight significant p-values.
  for (j in c(4,7,10,13)) {

    pvals <- suppressWarnings(as.numeric(ft$body$dataset[[j]]))
  sig_rows <- which(
  (!is.na(pvals) & pvals < 0.05) |
  ft$body$dataset[[j]] == "<0.001"
)

  ft <- ft %>%
    color(i = sig_rows, j = j, color = "red", part = "body")
}
  char_col <- names(ft$body$dataset)[1]
  age_row <- which(str_detect(ft$body$dataset[[char_col]], "^Age$"))
  rows_to_grey <- age_row:nrow(ft$body$dataset)
  ft = ft %>%
    color(i = rows_to_grey, color = "grey60", part = "body")
  
  ft = ft %>% add_header_lines("Adjusted Cox regression models") %>% bold(i = 1, part = "header")
  
}


```

## Build unadjusted Cox table function
```{r}
make_uni_cox_table <- function(data, mymarker) {

  tbls <- purrr::imap(
    outcomes,
    function(outcome, outcome_name) {

      tbl_uvregression(
        data = data,
        hide_n = TRUE,
        method = coxph,
        y = Surv(!!sym(outcome$time), !!sym(outcome$event)),
        include = all_of(mymarker),
        exponentiate = TRUE,
        pvalue_fun = format_p
      ) %>%
        add_global_p(keep = TRUE) %>%
        bold_p(t = 0.05) %>%
        bold_labels() %>%
        italicize_levels()
    }
  )

  ft <- tbl_merge(
    tbls = tbls,
    tab_spanner = paste0("**", purrr::map_chr(outcomes, "outcome_label"), "**")
  )

  ft = ft %>%
    as_flex_table() %>%
    set_table_properties(layout = "autofit") %>%
    padding(
      padding.top = 0, padding.bottom = 0,
      padding.left = 1, padding.right = 1
    ) %>%
    autofit() %>%
    vline(j = c(4, 7, 10), border = fp_border(color = "black", width = 1))

  for (j in c(4, 7, 10, 13)) {
    pvals <- suppressWarnings(as.numeric(ft$body$dataset[[j]]))
    sig_rows <- which(
      (!is.na(pvals) & pvals < 0.05) |
        ft$body$dataset[[j]] == "<0.001"
    )

    ft <- ft %>%
      color(i = sig_rows, j = j, color = "red", part = "body")
  }

  ft %>%
    add_header_lines("Unadjusted Cox regression models") %>%
    bold(i = 1, part = "header")
}
```


## Create event table function
```{r}
make_event_table <- function(data, marker, outcomes){

  purrr::imap_dfr(
    outcomes,
    function(outcome, outcome_name){

      data %>%
        dplyr::filter(!is.na(.data[[marker]])) %>%   # remove NA levels
        dplyr::group_by(.data[[marker]]) %>%
        dplyr::reframe(
          Events = sum(.data[[outcome$event]] == 1, na.rm = TRUE),
          N = dplyr::n()
        ) %>%
        dplyr::mutate(
          outcome = outcome$outcome_label,
          value = paste0(Events, " / ", N)
        ) %>%
        dplyr::select(!!marker := all_of(marker), outcome, value)
    }
  ) %>%
  tidyr::pivot_wider(
    names_from = outcome,
    values_from = value
  )
  
}
```

# Prognostic Performance: Uno's C-index
```{r}
# Uno's C-index is estimated using inverse-probability censoring weights
# (survival::concordance with timewt = "n/G2"; Uno et al., PMID: 21484848).

clinical_terms_all <- c(
  "age", "grade", "er", "her2_tpc_onco", "mpsiz", "pnod", "tmast", "trt"
)

clinical_terms_sub <- c(
  "age", "grade", "mpsiz", "pnod", "tmast", "trt"
)

make_uno_scenarios <- function(clinical_terms) {
  tibble::tibble(
    scenario = c(
      "Clinical",
      "TMEM doorway (tertiles)",
      "MenaINV (tertiles)",
      "MenaCalc (tertiles)",
      "Clinical + TMEM doorway (tertiles)",
      "Clinical + MenaINV (tertiles)",
      "Clinical + MenaCalc (tertiles)",
      "Clinical + TMEM doorway (tertiles) + MenaINV (tertiles)",
      "Clinical + TMEM doorway (tertiles) + MenaCalc (tertiles)",
      "Clinical + MenaINV (tertiles) + MenaCalc (tertiles)",
      "Clinical + TMEM doorway (tertiles) + MenaINV (tertiles) + MenaCalc (tertiles)"
    ),
    terms = list(
      clinical_terms,
      "tmem_tertile",
      "mena_inv_tertile",
      "mena_calc_tertile",
      c(clinical_terms, "tmem_tertile"),
      c(clinical_terms, "mena_inv_tertile"),
      c(clinical_terms, "mena_calc_tertile"),
      c(clinical_terms, "tmem_tertile", "mena_inv_tertile"),
      c(clinical_terms, "tmem_tertile", "mena_calc_tertile"),
      c(clinical_terms, "mena_inv_tertile", "mena_calc_tertile"),
      c(clinical_terms, "tmem_tertile", "mena_inv_tertile", "mena_calc_tertile")
    )
  )
}

build_uno_formula <- function(time, event, terms) {
  as.formula(
    paste0(
      "Surv(", time, ", ", event, ") ~ ",
      paste(terms, collapse = " + ")
    )
  )
}

fit_uno_cindex <- function(data, cohort, scenario, terms, outcome_name, outcome) {
  required_vars <- unique(c(outcome$time, outcome$event, terms))

  model_data <- data %>%
    dplyr::select(dplyr::all_of(required_vars)) %>%
    dplyr::filter(dplyr::if_all(dplyr::all_of(required_vars), ~ !is.na(.x))) %>%
    dplyr::filter(.data[[outcome$time]] > 0) %>%
    as.data.frame() %>%
    droplevels()

  n_events <- sum(model_data[[outcome$event]] == 1, na.rm = TRUE)
  follow_up <- ifelse(grepl("_5YS$", outcome_name), "5-year", "Overall follow-up")
  outcome_type <- sub("_5YS$", "", outcome_name)

  empty_result <- tibble::tibble(
    cohort = cohort,
    scenario = scenario,
    outcome = outcome_type,
    follow_up = follow_up,
    n = nrow(model_data),
    events = n_events,
    c_index = NA_real_,
    se = NA_real_,
    ci_lower = NA_real_,
    ci_upper = NA_real_,
    formula = NA_character_,
    note = NA_character_
  )

  if (nrow(model_data) < 2 || n_events == 0) {
    return(
      empty_result %>%
        dplyr::mutate(note = "Not enough observations or events to estimate C-index")
    )
  }

  f <- build_uno_formula(outcome$time, outcome$event, terms)
  formula_text <- paste(deparse(f), collapse = " ")

  fit <- tryCatch(
    survival::coxph(f, data = model_data, ties = "breslow", x = TRUE, y = TRUE),
    error = function(e) e
  )

  if (inherits(fit, "error")) {
    return(
      empty_result %>%
        dplyr::mutate(
          formula = formula_text,
          note = paste("Cox model failed:", conditionMessage(fit))
        )
    )
  }

  cfit <- tryCatch(
    survival::concordance(fit, timewt = "n/G2"),
    error = function(e) e
  )

  if (inherits(cfit, "error")) {
    return(
      empty_result %>%
        dplyr::mutate(
          formula = formula_text,
          note = paste("Concordance failed:", conditionMessage(cfit))
        )
    )
  }

  c_index <- as.numeric(cfit$concordance)[1]
  se <- sqrt(as.numeric(cfit$var)[1])

  tibble::tibble(
    cohort = cohort,
    scenario = scenario,
    outcome = outcome_type,
    follow_up = follow_up,
    n = fit$n,
    events = fit$nevent,
    c_index = c_index,
    se = se,
    ci_lower = pmax(0, c_index - 1.96 * se),
    ci_upper = pmin(1, c_index + 1.96 * se),
    formula = formula_text,
    note = NA_character_
  )
}

calculate_uno_cindex <- function(data, cohort, scenarios) {
  purrr::pmap_dfr(
    scenarios,
    function(scenario, terms) {
      purrr::imap_dfr(
        outcomes,
        ~ fit_uno_cindex(
          data = data,
          cohort = cohort,
          scenario = scenario,
          terms = terms,
          outcome_name = .y,
          outcome = .x
        )
      )
    }
  )
}

format_uno_estimate <- function(c_index, ci_lower, ci_upper, n, events) {
  dplyr::if_else(
    is.na(c_index),
    "",
    paste0(
      sprintf("%.3f", c_index),
      " (",
      sprintf("%.3f", ci_lower),
      "-",
      sprintf("%.3f", ci_upper),
      "); ",
      n,
      " / ",
      events
    )
  )
}

make_uno_cindex_table <- function(results, scenario_levels) {
  results %>%
    dplyr::mutate(
      scenario = factor(scenario, levels = scenario_levels),
      outcome_label = paste(outcome, follow_up),
      estimate = format_uno_estimate(c_index, ci_lower, ci_upper, n, events)
    ) %>%
    dplyr::select(Scenario = scenario, outcome_label, estimate) %>%
    tidyr::pivot_wider(
      names_from = outcome_label,
      values_from = estimate
    ) %>%
    dplyr::arrange(Scenario) %>%
    dplyr::mutate(Scenario = as.character(Scenario))
}

make_uno_summary_lines <- function(results, intro) {
  full_model_scenario <- paste(
    "Clinical + TMEM doorway (tertiles) + MenaINV (tertiles) +",
    "MenaCalc (tertiles)"
  )

  scenario_label <- function(x) {
    dplyr::case_when(
      x == full_model_scenario ~ "all three biomarkers",
      x == "Clinical + TMEM doorway (tertiles) + MenaINV (tertiles)" ~
        "Clinical + TMEM + MenaINV",
      x == "Clinical + TMEM doorway (tertiles) + MenaCalc (tertiles)" ~
        "Clinical + TMEM + MenaCalc",
      x == "Clinical + MenaINV (tertiles) + MenaCalc (tertiles)" ~
        "Clinical + MenaINV + MenaCalc",
      x == "Clinical + TMEM doorway (tertiles)" ~ "Clinical + TMEM",
      x == "Clinical + MenaINV (tertiles)" ~ "Clinical + MenaINV",
      x == "Clinical + MenaCalc (tertiles)" ~ "Clinical + MenaCalc",
      TRUE ~ x
    )
  }

  order_grid <- tibble::tibble(
    outcome = c("DDFS", "DDFS", "DFS", "DFS"),
    follow_up = c("Overall follow-up", "5-year", "Overall follow-up", "5-year"),
    line_label = c("DDFS overall", "DDFS 5-year", "DFS overall", "DFS 5-year")
  )

  clinical <- results %>%
    dplyr::filter(scenario == "Clinical") %>%
    dplyr::select(outcome, follow_up, clinical_c_index = c_index)

  best <- results %>%
    dplyr::group_by(outcome, follow_up) %>%
    dplyr::slice_max(c_index, n = 1, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(outcome, follow_up, best_scenario = scenario, best_c_index = c_index)

  summary_lines <- order_grid %>%
    dplyr::left_join(clinical, by = c("outcome", "follow_up")) %>%
    dplyr::left_join(best, by = c("outcome", "follow_up")) %>%
    dplyr::mutate(
      line = paste0(
        line_label,
        ": clinical ",
        sprintf("%.3f", clinical_c_index),
        "; best ",
        sprintf("%.3f", best_c_index),
        " with ",
        scenario_label(best_scenario)
      )
    ) %>%
    dplyr::pull(line)

  c(intro, summary_lines)
}

write_uno_cindex_sheet <- function(wb, sheet_name, table_data, summary_lines) {
  openxlsx::addWorksheet(wb, sheet_name)
  openxlsx::writeData(wb, sheet_name, table_data)

  header_style <- openxlsx::createStyle(textDecoration = "bold")
  summary_title_style <- openxlsx::createStyle(textDecoration = "bold", wrapText = TRUE)
  summary_style <- openxlsx::createStyle(wrapText = TRUE)

  openxlsx::addStyle(
    wb, sheet_name, header_style,
    rows = 1, cols = seq_len(ncol(table_data)), gridExpand = TRUE
  )
  openxlsx::setColWidths(
    wb, sheet_name,
    cols = seq_len(ncol(table_data)),
    widths = c(58, rep(34, ncol(table_data) - 1))
  )

  summary_start <- nrow(table_data) + 3
  openxlsx::writeData(
    wb, sheet_name,
    data.frame(summary = summary_lines),
    startRow = summary_start,
    colNames = FALSE
  )

  summary_rows <- summary_start + seq_along(summary_lines) - 1
  purrr::walk(
    summary_rows,
    ~ openxlsx::mergeCells(
      wb, sheet_name,
      cols = seq_len(ncol(table_data)),
      rows = .x
    )
  )
  openxlsx::addStyle(
    wb, sheet_name, summary_title_style,
    rows = summary_start,
    cols = 1,
    gridExpand = TRUE
  )
  openxlsx::addStyle(
    wb, sheet_name, summary_style,
    rows = summary_rows[-1],
    cols = 1,
    gridExpand = TRUE
  )
}

uno_scenarios_all <- make_uno_scenarios(clinical_terms_all)
uno_scenarios_sub <- make_uno_scenarios(clinical_terms_sub)

uno_cindex_all <- calculate_uno_cindex(
  data = df2,
  cohort = "All patients",
  scenarios = uno_scenarios_all
)

uno_cindex_sub <- calculate_uno_cindex(
  data = df2_sub,
  cohort = "ER+/HER2- subgroup",
  scenarios = uno_scenarios_sub
)

uno_cindex_results <- dplyr::bind_rows(uno_cindex_all, uno_cindex_sub)

uno_cindex_table_all <- make_uno_cindex_table(
  uno_cindex_all,
  uno_scenarios_all$scenario
)

uno_cindex_table_sub <- make_uno_cindex_table(
  uno_cindex_sub,
  uno_scenarios_sub$scenario
)

uno_summary_all <- make_uno_summary_lines(
  uno_cindex_all,
  "Across all patients, adding TMEM/Mena biomarkers gave modest improvement over clinical variables alone. The best gains were:"
)

uno_summary_sub <- make_uno_summary_lines(
  uno_cindex_sub,
  "In ER+/HER2- patients, C-index values were generally higher, and biomarker gains were slightly larger. The best gains were:"
)

uno_cindex_wb <- openxlsx::createWorkbook()
write_uno_cindex_sheet(
  uno_cindex_wb,
  "All patients",
  uno_cindex_table_all,
  uno_summary_all
)
write_uno_cindex_sheet(
  uno_cindex_wb,
  "ERpos HER2neg",
  uno_cindex_table_sub,
  uno_summary_sub
)
openxlsx::addWorksheet(uno_cindex_wb, "Numeric results")
openxlsx::writeData(uno_cindex_wb, "Numeric results", uno_cindex_results)
openxlsx::saveWorkbook(
  uno_cindex_wb,
  file.path(results_dir, "uno_cindex_results.xlsx"),
  overwrite = TRUE
)
```


# Main tables

## Table 1. Patient Characteristics by TMEM Doorway Score Tertiles
```{r}
set.seed(820)
table1_categorical_labels <- c(
  race2 = "Race",
  mense = "Menopausal status",
  grade = "Tumor grade",
  er = "Estrogen Receptor",
  pr = "Progesterone Receptor",
  her2_tpc_onco = "HER2 status",
  tmast = "Surgery",
  trt = "Treatment",
  tam = "Tamoxifen"
)

table1 = df %>% dplyr::select(tmem_tertile,age,race2, mense,grade, er,pr, `her2_tpc_onco`,mpsiz,pnod, tmast, trt, tam) %>% 
  tbl_summary(
    by = tmem_tertile, 
    missing = "ifany",
    missing_text = "(Missing)",
    digits = list(
     
      all_continuous() ~ 1,
      all_categorical() ~ c(0, 1)   # 0 digits for counts, 1 digit for percent
      
    ),
    label = list(
        age          ~  "Age (years)",
        race2         ~ "Race",
        mense         ~ "Menopausal status",
        grade         ~ "Tumor grade",
        er            ~ "Estrogen Receptor",
        pr            ~ "Progesterone Receptor",
        mpsiz         ~ "Tumor size (cm)",
        pnod          ~ "Positive lymph nodes",
        tmast         ~ "Surgery",
        trt           ~ "Treatment",
        tam           ~ "Tamoxifen"
      ),
    
    type = list(all_dichotomous() ~ "categorical",all_continuous() ~ "continuous2"),
    statistic = list(all_continuous() ~ c( 
                                           "{mean} ({sd})"),
                     all_categorical() ~ "{n} ({p})"))%>%  
  add_overall() %>% 
  add_p(
  test = list(
    all_continuous() ~ "aov",
    all_categorical() ~ "chisq.test"
  ),
  test.args = race_chisq_test_args,
  pvalue_fun = format_p_table1
) %>%   modify_header(label ~ "**Variables**") %>%
  modify_footnote(everything() ~ NA,all_stat_cols() ~ "Continuous variables were presented as mean (SD) and categorical variables were presented as N (%). P was calculated using ANOVA for continuous variables and chi-square for categorical variables; Monte Carlo simulation was used for Race.") %>% 
  bold_labels()%>%  bold_p(t = 0.05) %>% italicize_levels()

table1 = as_flex_table(table1)%>% set_table_properties(layout = "autofit") %>% padding(padding.top = 0, padding.bottom = 0, padding.left = 1, padding.right = 1) %>% autofit()
table1
```


## Save main analysis tables
```{r}
mymarkers = main_markers

doc = new_results_doc()

doc = doc %>% body_add_par("Table 1. Baseline Characteristics by TMEM Doorway Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(table1)

doc = doc %>% body_add_par("Table 2 / Supplementary Table S4. Adjusted Cox Model Outputs Among All Patients", style = "heading 1")

#####all patients
for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("All patients - ",mymarker), style = "heading 2") 
  
  if (mymarker %in% tertile_markers){
  event_table <- make_event_table(df2, mymarker, outcomes)
  event_table_ft <- flextable(event_table) %>%
  autofit() %>% add_header_lines("Number of Events / Number at Risk by Marker Level")%>% bold(i = 1, part = "header")
    
    doc = doc %>% body_add_flextable(event_table_ft) 
  }

  doc = doc %>% 
  body_add_flextable(make_uni_cox_table(df2, mymarker)) %>% 
  body_add_par("", style = "Normal") %>% 
  body_add_flextable(make_cox_table(df2, mymarker,mycovars)) 
  
}

#### subgroup
doc = doc %>% body_add_par("Tables 4 and 5. Adjusted Cox Model Outputs in the ER+/HER2- Subgroup", style = "heading 1")

for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("ER+/HER2- subgroup - ",mymarker), style = "heading 2") 
  
  if (mymarker %in% tertile_markers){
  event_table <- make_event_table(df2_sub, mymarker, outcomes)
  event_table_ft <- flextable(event_table) %>%
  autofit() %>% add_header_lines("Number of Events / Number at Risk by Marker Level")%>% bold(i = 1, part = "header")
  doc = doc %>% body_add_flextable(event_table_ft) 
  }
  doc = doc %>% 
  body_add_flextable(make_uni_cox_table(df2_sub, mymarker)) %>% 
  body_add_par("", style = "Normal") %>% 
  body_add_flextable(make_cox_table(df2_sub, mymarker, subcovars)) 
  
}

path = file.path(results_dir, "main_analysis_tables.docx")
print(doc, path)
```


# Supplementary tables

## Supp Table 2. Patient Characteristics by MenaCalc Score Tertiles
```{r}
set.seed(820)
stable2 = df %>% dplyr::select(mena_calc_tertile,age,race2, mense,grade, er,pr, `her2_tpc_onco`,mpsiz,pnod, tmast, trt, tam) %>% 
  tbl_summary(
    by = mena_calc_tertile, 
    missing = "ifany",
    missing_text = "(Missing)",
    digits = list(
     
      all_continuous() ~ 1,
      all_categorical() ~ c(0, 1)    # 0 digits for counts, 1 digit for percent
    ),
    label = list(
        age          ~  "Age (years)",
        race2         ~ "Race",
        mense         ~ "Menopausal status",
        grade         ~ "Tumor grade",
        er            ~ "Estrogen Receptor",
        pr            ~ "Progesterone Receptor",
        mpsiz         ~ "Tumor size (cm)",
        pnod          ~ "Positive lymph nodes",
        tmast         ~ "Surgery",
        trt           ~ "Treatment",
        tam           ~ "Tamoxifen"
      ),
    
    type = list(all_dichotomous() ~ "categorical",all_continuous() ~ "continuous2"),
    statistic = list(all_continuous() ~ c( 
                                           "{mean} ({sd})"),
                     all_categorical() ~ "{n} ({p})"))%>%  
  add_overall() %>% 
  add_p(
  test = list(
    all_continuous() ~ "aov",
    all_categorical() ~ "chisq.test"
  ),
  test.args = race_chisq_test_args,
  pvalue_fun = format_p_table1
) %>%   modify_header(label ~ "**Variables**") %>%
  modify_footnote(everything() ~ NA,all_stat_cols() ~ "Continuous variables were presented as mean (SD) and categorical variables were presented as N (%). P was calculated using ANOVA for continuous variables and chi-square for categorical variables; Monte Carlo simulation was used for Race.") %>% 
  bold_labels()%>%  bold_p(t = 0.05) %>% italicize_levels()

stable2 = as_flex_table(stable2)%>% set_table_properties(layout = "autofit") %>% padding(padding.top = 0, padding.bottom = 0, padding.left = 1, padding.right = 1) %>% autofit()
stable2

```

## Supp Table 3. Patient Characteristics by MenaINV Tertiles
```{r}
set.seed(820)
stable3 = df %>% dplyr::select(mena_inv_tertile,age,race2, mense,grade, er,pr, `her2_tpc_onco`,mpsiz,pnod, tmast, trt, tam) %>% 
  tbl_summary(
    by = mena_inv_tertile, 
    missing = "ifany",
    missing_text = "(Missing)",
    digits = list(
     
      all_continuous() ~ 1,
      all_categorical() ~ c(0, 1)    # 0 digits for counts, 1 digit for percent
       
    ),
    label = list(
        age          ~  "Age (years)",
        race2         ~ "Race",
        mense         ~ "Menopausal status",
        grade         ~ "Tumor grade",
        er            ~ "Estrogen Receptor",
        pr            ~ "Progesterone Receptor",
        mpsiz         ~ "Tumor size (cm)",
        pnod          ~ "Positive lymph nodes",
        tmast         ~ "Surgery",
        trt           ~ "Treatment",
        tam           ~ "Tamoxifen"
      ),
    
    type = list(all_dichotomous() ~ "categorical",all_continuous() ~ "continuous2"),
    statistic = list(all_continuous() ~ c( 
                                           "{mean} ({sd})"),
                     all_categorical() ~ "{n} ({p})"))%>%  
  add_overall() %>% 
  add_p(
  test = list(
    all_continuous() ~ "aov",
    all_categorical() ~ "chisq.test"
  ),
  test.args = race_chisq_test_args,
  pvalue_fun = format_p_table1
) %>%   modify_header(label ~ "**Variables**") %>%
  modify_footnote(everything() ~ NA,all_stat_cols() ~ "Continuous variables were presented as mean (SD) and categorical variables were presented as N (%). P was calculated using ANOVA for continuous variables and chi-square for categorical variables; Monte Carlo simulation was used for Race.") %>% 
  bold_labels()%>%  bold_p(t = 0.05) %>% italicize_levels()

stable3 = as_flex_table(stable3)%>% set_table_properties(layout = "autofit") %>% padding(padding.top = 0, padding.bottom = 0, padding.left = 1, padding.right = 1) %>% autofit()
stable3
```
## Supp Table 4. Spearman Correlations (rs) Within TMEM, MenaCalc, and MenaINV in All Patients and the ER+HER2- Subgroup

```{r}
vars <- c("tmem_score","mena_calc_score","mena_inv_score")

spearman_results <- combn(vars, 2, simplify = FALSE) %>%
  map_dfr(~{
    test <- cor.test(
      df[[.x[1]]],
      df[[.x[2]]],
      method = "spearman",
      exact = FALSE   # recommended for larger samples / ties
    )
    tidy(test) %>%
      mutate(
        var1 = .x[1],
        var2 = .x[2],
        method = "Spearman"
      )
  })

spearman_results = spearman_results  %>% mutate(p.value = ifelse(
      p.value < 0.001,
      "<0.001",
      sprintf("%.3f", p.value))) %>% mutate(rho = round(estimate, 3))%>% select(var1, var2,rho,everything()) %>% select(-estimate, -statistic,-method,-alternative)
spearman_results = flextable(spearman_results)
spearman_results

spearman_results_sub <- combn(vars, 2, simplify = FALSE) %>%
  map_dfr(~{
    test <- cor.test(
      df_sub[[.x[1]]],
      df_sub[[.x[2]]],
      method = "spearman",
      exact = FALSE   # recommended for larger samples / ties
    )
    tidy(test) %>%
      mutate(
        var1 = .x[1],
        var2 = .x[2],
        method = "Spearman"
      )
  })

spearman_results_sub = spearman_results_sub  %>% mutate(p.value = ifelse(
      p.value < 0.001,
      "<0.001",
      sprintf("%.3f", p.value))) %>% mutate(rho = round(estimate, 3))%>% select(var1, var2,rho,everything()) %>% select(-estimate, -statistic,-method,-alternative)
spearman_results_sub = flextable(spearman_results_sub)
spearman_results_sub
```

## Run numeric-marker p-trend models
```{r}
mymarkers = numeric_mutual_markers
doc = new_results_doc()

doc = doc %>% body_add_par("Table 3 / Supplementary Table S5 p trend in All Patients", style = "heading 1")

#####all patients
for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("All patients - ",mymarker), style = "heading 2") 

  doc = doc %>% 
  body_add_flextable(make_cox_table(df2, mymarker,mycovars)) 
  
}

#### subgroup
doc = doc %>% body_add_par("Table 3 / Supplementary Table S5 p trend in the ER+/HER2- Subgroup", style = "heading 1")

for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("ER+/HER2- subgroup - ",mymarker), style = "heading 2") 
  
  doc = doc %>% 
  body_add_flextable(make_cox_table(df2_sub, mymarker, subcovars)) 
} 

path = file.path(results_dir, "table3_supplementary_tables5p_trend_tables.docx")
print(doc, path)

```


## Supplementary Table S6. Treatment-arm interaction by tertile increase
```{r}
format_hr_ci <- function(estimate, se) {
  sprintf(
    "%.2f (%.2f, %.2f)",
    exp(estimate),
    exp(estimate - 1.96 * se),
    exp(estimate + 1.96 * se)
  )
}

get_marker_trt_interaction <- function(data, marker_num, time, event) {
  covars <- " + age + grade + pnod + er + mpsiz + tmast + her2_tpc_onco"
  f_full <- as.formula(paste0("Surv(", time, ", ", event, ") ~ ", marker_num, " * trt", covars))
  f_reduced <- as.formula(paste0("Surv(", time, ", ", event, ") ~ ", marker_num, " + trt", covars))

  fit <- coxph(f_full, data = data, ties = "breslow", x = TRUE, y = TRUE)
  reduced_fit <- coxph(f_reduced, data = data, ties = "breslow", x = TRUE, y = TRUE)

  beta <- coef(fit)
  vc <- vcov(fit)
  interaction_term <- grep(
    paste0("(^", marker_num, ":trt)|(^trt.*:", marker_num, "$)"),
    names(beta),
    value = TRUE
  )

  if (length(interaction_term) != 1) {
    stop("Could not uniquely identify treatment interaction term for ", marker_num)
  }

  ac_est <- beta[[marker_num]]
  ac_se <- sqrt(vc[marker_num, marker_num])
  acp_est <- beta[[marker_num]] + beta[[interaction_term]]
  acp_se <- sqrt(
    vc[marker_num, marker_num] +
      vc[interaction_term, interaction_term] +
      2 * vc[marker_num, interaction_term]
  )
  p_interaction <- anova(reduced_fit, fit, test = "LRT")[2, "Pr(>|Chi|)"]

  tibble(
    Treatment = c("AC", "AC + PTX"),
    HR = c(format_hr_ci(ac_est, ac_se), format_hr_ci(acp_est, acp_se)),
    `P interaction` = c(format_p(p_interaction), "")
  )
}

s6_markers <- tribble(
  ~marker_num,     ~marker_label,
  "tmem_num",      "TMEM",
  "mena_calc_num", "MenaCalc",
  "mena_inv_num",  "MenaINV"
)

s6_outcomes <- tribble(
  ~Outcome,         ~time,       ~event,
  "All-year DDFS",  "tddfs",     "ddfs",
  "5-year DDFS",    "tddfs_5ys", "ddfs_5ys"
)

make_s6_outcome_rows <- function(Outcome, time, event) {
  marker_tables <- pmap(
    s6_markers,
    function(marker_num, marker_label) {
      get_marker_trt_interaction(df2, marker_num, time, event) %>%
        rename(
          !!paste0(marker_label, " HR (95% CI)") := HR,
          !!paste0(marker_label, " P interaction") := `P interaction`
        )
    }
  )

  reduce(marker_tables, left_join, by = "Treatment") %>%
    mutate(Outcome = Outcome, .before = Treatment)
}

stable6_marker_rows <- pmap_dfr(s6_outcomes, make_s6_outcome_rows)

stable6 <- flextable(stable6_marker_rows) %>%
  add_header_lines(
    "Supplementary Table S6. Adjusted HRs per Tertile Increase and Interaction P Values by Treatment Arm"
  ) %>%
  add_footer_lines(
    "Adjusted for age, tumor grade, ER, HER2 status, tumor size, positive lymph nodes, surgery, treatment, and marker-by-treatment interaction. P interaction values were assessed using likelihood-ratio tests. Each exposure was evaluated as an ordinal variable, and each level (tertile) corresponded to the equivalent integer (from 1 through 3)."
  ) %>%
  theme_booktabs() %>%
  autofit()

stable6
```

## Save supplementary analysis tables
```{r}
mymarkers = tertile_mutual_markers
doc = new_results_doc()

doc = doc %>% body_add_par("Supplementary Table S2. Baseline Characteristics by MenaCalc Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(stable2)

doc = doc %>% body_add_par("Supplementary Table S3. Baseline Characteristics by MenaINV Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(stable3)

doc = doc %>% body_add_par("Additional Analysis. Spearman Correlations within TMEM, MenaCalc, and MenaINV", style = "heading 1")
doc = doc %>% body_add_par("All Patients:", style = "Normal")
doc = doc %>% body_add_flextable(spearman_results)
doc = doc %>% body_add_par("ER+HER2- Subgroup:", style = "Normal")
doc = doc %>% body_add_flextable(spearman_results_sub)

doc = doc %>% body_add_par("Table 3 / Supplementary Table S5. Mutually Adjusted Associations in All Patients", style = "heading 1")

for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("All patients - ",mymarker), style = "heading 2") 

  doc = doc %>% 
  body_add_flextable(make_cox_table(df2, mymarker,mycovars)) 
  
}

doc = doc %>% body_add_par("Table 3 / Supplementary Table S5. Mutually Adjusted Associations in the ER+/HER2- Subgroup", style = "heading 1")

for (mymarker in mymarkers){

  doc = doc %>% body_add_par(paste0("ER+/HER2- subgroup - ",mymarker), style = "heading 2") 
  
  doc = doc %>% 
  body_add_flextable(make_cox_table(df2_sub, mymarker, subcovars)) 
} 

doc = doc %>% body_add_par("Supplementary Table S6. Treatment Arm Interaction by Tertile Increase", style = "heading 1")
doc = doc %>% body_add_flextable(stable6)

path = file.path(results_dir, "supplementary_analysis_tables.docx")
print(doc, path)
```

# Kaplan-Meier curves

### Function to produce one KM plot

```{r}
kmf <- function(my_var,mydata, my_time, my_event, my_title, my_subtitle) {
  if (my_var == "tmem_tertile"){
    my_legend = c("Low", "Mid", "High")
    my_legend_title = "TMEM Doorway Score"
  } else if (my_var == "mena_calc_tertile"){
    my_legend = c("Low", "Mid", "High")
    my_legend_title = expression(Mena^{calc}~"Score")
  } else if (my_var == "mena_inv_tertile"){
    my_legend = c("Low", "Mid", "High")
    my_legend_title = expression(Mena^{INV}~"Score")
  } 
  
  # Create the formula using reformulate() for reliable variable evaluation.
  surv_formula <- reformulate(my_var, 
                            response = paste0("Surv(", my_time, ", ", my_event, ")"))
  
  sd <- survival::survdiff(surv_formula, data = mydata)
  pval_raw <- 1 - stats::pchisq(sd$chisq, df = length(sd$n) - 1)
  
  pval_label <- ifelse(is.na(pval_raw), "", ifelse(pval_raw < 0.001, "P < 0.001", ifelse(pval_raw < 0.01, paste0("P = ", sprintf("%.3f", pval_raw)), paste0("P = ", sprintf("%.2f", pval_raw)))))


  # Fit the model with explicit data evaluation for plotting.
  fit <- eval(bquote(
    survfit(.(surv_formula), data = mydata,)
  ))
  
  p1 <- ggsurvplot(
    fit,
    data = mydata,
    title = paste("Kaplan-Meier Curve of", my_title),
    subtitle = my_subtitle,
    legend.title = my_legend_title,
    legend.labs = my_legend,
    font.title = c(12, "bold", "black"),
    font.subtitle = c(10, "bold.italic", "black"),
    font.x = c(10, "italic", "black"),
    font.y = c(10, "italic", "black"),
    fontsize = 3,
    font.tickslab = c(9, "plain", "black"),
    risk.table.height = 0.25,
    pval = pval_label, 
    conf.int = FALSE,
    risk.table = TRUE,
    risk.table.col = "strata",
    linetype = "strata",
    test.for.trend = FALSE,
    xlab = "Time (years)",
    ylab = "Survival",
    risk.table.y.text = FALSE,
    font.legend = list(size = 10, color = "black"),
    xscale = "m_y",          # months → years
    break.time.by = 12,      # tick every year
    ggtheme = theme_classic2(base_size = 10)
  )
  
  p1$table <- p1$table + 
    theme(plot.title = element_text(size = 9))
  
  return(p1)
}

```
### Function to generate and save multiple KM plots
```{r}

generate_km_plots <- function(mydata, primary_var, file_tag) {
  
  km.ddfs <- kmf(primary_var, mydata, "tddfs", "ddfs", "Distant Disease Free Survival", "Without truncation")
  km.ddfs.5ys <- kmf(primary_var, mydata, "tddfs_5ys", "ddfs_5ys", "Distant Disease Free Survival", "With 5-year truncation")
  
  km.dfs <- kmf(primary_var, mydata, "tdfs", "dfs", "Disease Free Survival", "Without truncation")
  km.dfs.5ys <- kmf(primary_var, mydata, "tdfs_5ys", "dfs_5ys", "Disease Free Survival", "With 5-year truncation")

  cohort_label <- dplyr::recode(
    file_tag,
    "all_sample" = "all_patients",
    "her2_er_subgroup" = "er_her2",
    .default = file_tag
  )
  marker_label <- dplyr::recode(
    primary_var,
    "tmem_tertile" = "tmem_tertile",
    "mena_calc_tertile" = "menacalc_tertile",
    "mena_inv_tertile" = "menainv_tertile",
    .default = primary_var
  )

  pdf(file.path(km_dir, paste0(cohort_label, "_", marker_label, ".pdf")), width = 10, height = 5)

  arrange_ggsurvplots(list(km.ddfs, km.ddfs.5ys), ncol = 2, nrow = 1)
  arrange_ggsurvplots(list(km.dfs, km.dfs.5ys), ncol = 2, nrow = 1)

  dev.off()
}
```

### Produce and save KM curves
```{r}
for (primary_var in tertile_markers){
 generate_km_plots(df, primary_var, "all_sample")
}

for (primary_var in tertile_markers){
 generate_km_plots(df_sub, primary_var, "her2_er_subgroup")
}
```


```{r}
# Save distribution tables.
doc = new_results_doc()

doc = doc %>% body_add_par("Table 1. Baseline Characteristics by TMEM Doorway Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(table1)

doc = doc %>% body_add_par("Supplementary Table S2. Baseline Characteristics by MenaCalc Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(stable2)

doc = doc %>% body_add_par("Supplementary Table S3. Baseline Characteristics by MenaINV Score Tertiles", style = "heading 1")
doc = doc %>% body_add_flextable(stable3)

path = file.path(results_dir, "1. Main table 1 Supplementary Table S2 S3.docx")
print(doc, path)
```

# Additional code 
## Reproduce manuscript Table 4: ER+/HER2- subgroup
```{r}
# This final standalone chunk does not modify the primary-analysis variables.
# Tertile cutoffs are calculated among all patients in df.
table4_cutpoints <- list(
  tmem = quantile(df$tmem_score, c(1 / 3, 2 / 3), na.rm = TRUE),
  mena_calc = quantile(df$mena_calc_score, c(1 / 3, 2 / 3), na.rm = TRUE),
  mena_inv = quantile(df$mena_inv_score, c(1 / 3, 2 / 3), na.rm = TRUE)
)

recurrence_summary_text <- function(data, cohort_label) {
  recurrence_values <- data$recurrence

  sprintf(
    "%s: mean %.2f, median %.2f, range %.2f-%.2f; missing %d of %d.",
    cohort_label,
    mean(recurrence_values, na.rm = TRUE),
    median(recurrence_values, na.rm = TRUE),
    min(recurrence_values, na.rm = TRUE),
    max(recurrence_values, na.rm = TRUE),
    sum(is.na(recurrence_values)),
    length(recurrence_values)
  )
}

table4_df_recurrence_summary <- recurrence_summary_text(
  df,
  "Full cohort (df) recurrence score"
)

table4_sub_recurrence_summary <- recurrence_summary_text(
  df_sub,
  "ER+/HER2- subgroup (df_sub) recurrence score"
)

table4_recurrence_imputation_note <- sprintf(
  paste0(
    "For Cox regression analyses in df2, missing recurrence scores were ",
    "replaced with the full-cohort median (%.2f)."
  ),
  median(df$recurrence, na.rm = TRUE)
)

table4_df <- df2_sub %>%
  mutate(
    table4_tmem = cut(
      tmem_score, c(-Inf, table4_cutpoints$tmem, Inf),
      labels = c("Low", "Mid", "High"), right = FALSE
    ),
    table4_mena_calc = cut(
      mena_calc_score, c(-Inf, table4_cutpoints$mena_calc, Inf),
      labels = c("Low", "Mid", "High"), right = FALSE
    ),
    table4_mena_inv = cut(
      mena_inv_score, c(-Inf, table4_cutpoints$mena_inv, Inf),
      labels = c("Low", "Mid", "High"), right = FALSE
    )
  )

table4_markers <- tribble(
  ~marker,            ~score,             ~label,
  "table4_tmem",      "tmem_score",       "TMEM",
  "table4_mena_calc", "mena_calc_score",  "MenaCalc",
  "table4_mena_inv",  "mena_inv_score",   "MenaINV"
)

table4_outcomes <- tribble(
  ~outcome, ~time,   ~event,
  "DDFS",   "tddfs", "ddfs",
  "DFS",    "tdfs",  "dfs"
)

table4_model_results <- function(data, marker, score, time, event, covars) {
  categorical_fit <- coxph(
    as.formula(paste0("Surv(", time, ", ", event, ") ~ ", marker, covars)),
    data = data,
    ties = "breslow",
    model = TRUE
  )

  trend_fit <- coxph(
    as.formula(paste0("Surv(", time, ", ", event, ") ~ ", score, covars)),
    data = data,
    ties = "breslow"
  )

  estimates <- broom::tidy(
    categorical_fit,
    exponentiate = TRUE,
    conf.int = TRUE
  )

  hr <- c(
    "1.00",
    map_chr(levels(data[[marker]])[-1], function(level) {
      result <- estimates %>% filter(term == paste0(marker, level))
      sprintf(
        "%.2f (%.2f, %.2f)",
        result$estimate, result$conf.low, result$conf.high
      )
    })
  )

  list(
    hr = hr,
    p_trend = format_p(
      car::Anova(
        trend_fit,
        type = 3,
        test.statistic = "LR"
      )[score, "Pr(>Chisq)"]
    )
  )
}

table4_rows <- pmap_dfr(table4_outcomes, function(outcome, time, event) {
  marker_rows <- pmap_dfr(
    table4_markers,
    function(marker, score, label) {
      counts <- map_chr(levels(table4_df[[marker]]), function(level) {
        level_data <- table4_df %>% filter(.data[[marker]] == level)
        paste0(
          sum(level_data[[event]] == 1, na.rm = TRUE),
          " / ",
          nrow(level_data)
        )
      })

      adjusted <- table4_model_results(
        table4_df, marker, score, time, event, subcovars
      )

      tibble(
        Characteristic = c(label, "#Event/#At Risk", "Adjusted HR"),
        T1 = c("", counts[1], adjusted$hr[1]),
        T2 = c("", counts[2], adjusted$hr[2]),
        T3 = c("", counts[3], adjusted$hr[3]),
        `P trend` = c("", "", adjusted$p_trend)
      )
    }
  )

  bind_rows(
    tibble(
      Characteristic = outcome,
      T1 = "", T2 = "", T3 = "", `P trend` = ""
    ),
    marker_rows
  )
})

table4_ft <- flextable(table4_rows) %>%
  set_header_labels(Characteristic = "") %>%
  add_header_lines(
    "Table 4. Adjusted Hazard Ratios (95% Confidence Intervals) for the Associations of TMEM Doorway Score, MenaCalc Score, and MenaINV Score With DDFS and DFS in the ER+/HER2- Subgroup"
  ) %>%
  theme_booktabs() %>%
  bold(
    i = which(
      table4_rows$Characteristic %in%
        c(table4_outcomes$outcome, table4_markers$label)
    ),
    j = 1,
    part = "body"
  ) %>%
  align(j = 1, align = "left", part = "all") %>%
  align(j = 2:5, align = "center", part = "all") %>%
  autofit() %>%
  add_footer_lines(
    "Adjusted for age (years), tumor grade, tumor size (cm), positive lymph nodes, surgery, and treatment; recurrence score was not included."
  )

table4_ft

```


## Table 4 sensitivity analysis: additionally adjusted for recurrence
```{r}
table4_recurrence_covars <- paste0(subcovars, " + recurrence")

table4_recurrence_rows <- pmap_dfr(
  table4_outcomes,
  function(outcome, time, event) {
    marker_rows <- pmap_dfr(
      table4_markers,
      function(marker, score, label) {
        counts <- map_chr(levels(table4_df[[marker]]), function(level) {
          level_data <- table4_df %>%
            filter(.data[[marker]] == level)

          paste0(
            sum(level_data[[event]] == 1, na.rm = TRUE),
            " / ",
            nrow(level_data)
          )
        })

        adjusted <- table4_model_results(
          table4_df,
          marker,
          score,
          time,
          event,
          table4_recurrence_covars
        )

        tibble(
          Characteristic = c(label, "#Event/#At Risk", "Adjusted HR"),
          T1 = c("", counts[1], adjusted$hr[1]),
          T2 = c("", counts[2], adjusted$hr[2]),
          T3 = c("", counts[3], adjusted$hr[3]),
          `P trend` = c("", "", adjusted$p_trend)
        )
      }
    )

    bind_rows(
      tibble(
        Characteristic = outcome,
        T1 = "", T2 = "", T3 = "", `P trend` = ""
      ),
      marker_rows
    )
  }
)

table4_recurrence_ft <- flextable(table4_recurrence_rows) %>%
  set_header_labels(Characteristic = "") %>%
  add_header_lines(
    "Table 4. Adjusted Hazard Ratios (95% Confidence Intervals) for the Associations of TMEM Doorway Score, MenaCalc Score, and MenaINV Score With DDFS and DFS in the ER+/HER2- Subgroup, Additionally Adjusted for Recurrence"
  ) %>%
  theme_booktabs() %>%
  bold(
    i = which(
      table4_recurrence_rows$Characteristic %in%
        c(table4_outcomes$outcome, table4_markers$label)
    ),
    j = 1,
    part = "body"
  ) %>%
  align(j = 1, align = "left", part = "all") %>%
  align(j = 2:5, align = "center", part = "all") %>%
  autofit() %>%
  add_footer_lines(
    "Adjusted for age (years), tumor grade, tumor size (cm), positive lymph nodes, surgery, treatment, and recurrence score. Missing recurrence scores were replaced with the full-cohort median."
  )

table4_recurrence_ft

table4_recurrence_doc <- new_results_doc() %>%
  body_add_par(table4_df_recurrence_summary, style = "Normal") %>%
  body_add_par(table4_sub_recurrence_summary, style = "Normal") %>%
  body_add_par(table4_recurrence_imputation_note, style = "Normal") %>%
  body_add_flextable(table4_recurrence_ft)

print(
  table4_recurrence_doc,
  file.path(
    results_dir,
    "table4_erpos_her2neg_adjusted_for_recurrence.docx"
  )
)
```
